Automated topic naming to support cross-project analysis of software maintenance activities

### 2011-01-23 msr/msr2011.tex
made more subsubsections

@@

\noindent \textbf{The high-level labels} --- To facilitate cross-project comparison, we used a taxonomy of NFRs. This taxonomy is based on the ISO quality model, ISO9126~\cite{iso9126}. 
ISO9126 describes six high-level non-functional requirements (NFRs), maintainability, functionality, portability, efficiency, usability, and reliability. 
There is some debate about the terms in this model. 
However, ISO9126 is ``an international standard and thus provides an internationally accepted terminology for software quality \cite[p. 58]{Boegh2008},'' that is sufficient for the purposes of this research.  
We claim that these NFRs are maintenance concerns (to varying degrees) in all software projects, and are therefore well-suited to comparison between projects.

\noindent \textbf{Creating a validation corpus} --- To evaluate both unsupervised and supervised classification, we created a validation set of manually labelled topics. For MySQL 3.23 and MaxDB 7.500, we annotated each extracted topic in each period with the same NFR labels as described in the preceding paragraph. 
We looked at each period's topics, and assessed what the data --- consisting of the frequency-weighted word lists and messages --- suggested was the topic for that period. 
We were able to pinpoint the appropriate label using auxiliary information as well, such as the actual revisions and files that were related to the topic being annotated.
For example, for the MaxDB topic consisting of a message ``exit() only used in non NPTL LINUX Versions'', we tagged that topic \emph{portability}. 
%\todo[inline]{this confused people. Is it truly unsupervised? Yes, but it needs to be clearer.}
%We compared against this data-set, but we also used the data-set for our supervised machine learning based topic classification. 
% A topic may have more than one matching keyword. % ok what?
%\todo[inline]{inter-rater reliability }
% \todo[inline]{There should be a section here which is about manual annotation. It is really important and it is required by both analyses. Plus we can tell them how.}

@@

\todo[inline]{NEIL: abram, i got rid of the letter-grade thing because I'm worried people latch on to the idea we are C- students. I feel anything better than random is our criteria here}
ROC values provide a score %, similar to school letter-grades (A is 0.9, C is 0.6),
 reflecting how well a particular learner performed for the given data. 
ROC maps to the more familiar concepts of precision/sensitivity and recall/specificity: it plots the true positive rate (sensitivity) versus the false positive rate (1 - specificity). 
A perfect learner has a ROC value of 1.0, reflecting perfect recall and precision. 
A ROC result of 0.5 would be equivalent to a random learner (that is, issuing as many false positives as true positives). 
The %area under the -- NE confusing to have area again ... and not importantana
ROC of a classifier is equivalent to the probability that the classifier will rank a randomly chosen positive instance higher than a randomly chosen negative instance.

We consider our labelling classifiers acceptable if they outperform a random classifier (0.5). 
\subsection{Unsupervised labelling}
\label{sec:unsuplabelling}

@@

\noindent\textbf{Generating word lists} --- In order to automatically label each topic with an NFR from ISO9126, we associate each NFR with a list
of keywords (`word-lists', in our parlance). These word-lists were determined a priori and were not extracted from the projects themselves.
We intersected the words of the topics and the words of our word-lists.
We `labeled' a topic if any of its words matched any of the word-list's words.
A topic could match more than one NFR.
We used several different sets of word-lists for comparison, which we refer to as \textsf{exp1, exp2, exp3} in the text which follows.

@@

\todo[inline]{NEIL: this could be a subsubsection}
%\todo[inline]{* the hierarchy of tags was not clear}
\noindent \textbf{Generating labels automatically} --- Using our three word-lists (\textsf{exp1}, \textsf{exp2}, \textsf{exp3}), we labeled our topics with an NFR where there was a match between a word in the list and the same word somewhere in the distribution of words that constitute the topic.
A \emph{named topic} is a topic with a matching NFR label. Unnamed topics occur where this no such match, which may indicate either a lack of precision, or simply that this topic is not associated with non-functional requirements.
All experiments were run on MaxDB 7.500 and MySQL 3.23 data. LDA
extracted 20 topics per period for each.
% Is this good enough
This labelling is unsupervised because the corpus is not derived from 
the project being labelled, and we did not label the project's topics
ourselves for a training set. The theory behind this technique is that
because most software often addresses similar issues, we can use the
domain knowledge of software to label relevant topics.
%\todo[inline]{Neil does this explain the unsupervised aspect well enough?}

@@

\noindent \textbf{Analysis of the unsupervised labelling} --- %Based on the labels, and our manual topic labelling, 
%we compared the results of the unsupervised word matching approach. 
For each quality we assessed whether unsupervised labels matched the manual annotations. %To be clear, the manual annotations were not used to train the labeling process.
As described in \ref{sec:wordlist} we used both ROC and F-1 measures to evaluate the performance of the classification.
Figure \ref{fig:maxdb-unsup-results} shows our ROC results for MaxDB and MySQL. We describe F-1 results in the text below.

@@

\textbf{Analysis of the supervised labelling} -- 
% \label{sec:suplabelling}
Because our data-set was of word counts we expected Bayesian techniques, often used in spam filtering, to perform well. 
We tried other learners that WEKA~\cite{weka09} provides: rule learners, decision tree learners, vector space learners, and support vector machines.  
Figure \ref{fig:best-learn-per-tag} shows the performance of the best performing learner per label. 
 We considered the best learner for a label to be the learner which
 had the highest ROC value for that label. 
The best learner is important because one uses a single learner per label.
%Figure \ref{fig:best-learn-per-tag} uses the ZeroR learner as a baseline, a common practice in machine learning,
%as ZeroR naively chooses the largest category all of the time. 
%ZeroR occasionally outperforms other learners; for labels which are not as common, this is to be expected because any miscategorization will hurt accuracy. 
%This is why we use ROC values, instead of accuracy, as they can better represent performance on labels which are not applicable to the majority of samples.

@@

\section{Understanding software maintenance activities} 
\label{sec:analysis}
% \todo[inline]{talk about this topic in one section}
% \todo[inline]{reference project name Nomen more}
Since our ROC values produced acceptable results, we now turn to \emph{applying} those techniques to understand software maintenance activities. We evaluated two research questions:
\begin{enumerate}
\item Do label frequencies change over time? Is a certain quality of more interest at one point in the life-cycle than some other? 
\item  Do the different projects differ in their relative topic interest? Is a particular quality more important to one project than the other projects?  
\end{enumerate}

% There are several possible applications of \textsf{Nomen}'s techniques. 
% For outside users of the software, \textsf{Nomen} can provide insight into that company's development practices. 
% Are they focused on usability, for example?
% Within the organization \textsf{Nomen} acts as a historical sanity check. 
% Confirmation bias is the difference between what people think happened and what actually happened. 
% Some software releases, for instance, are poorly received since they are quite buggy. 
% \textsf{Nomen} would help explain why this was the case, since it can identify NFRs which are dominating developer attention.
\todo[inline]{does this section use supervised learning, and if so what ones?}

\subsection{Timeline analysis}
Figures \ref{fig:mysql-timeline} and \ref{fig:maxdb-timeline} show the temporal patterns of label frequencies. 
There are two measures represented. 
One, the relative frequency, shown in the grey boxes, represents the number of labels with that NFR in that period, 
relative to the maximum number of topics assigned to that NFR. 
The second, absolute frequency, compares the number of topics labelled with that NFR per period 
relative to the maximum number of labelled topics overall, for that project. 
The topmost row in each diagram is reserved for historical events of note. 
We refer to these occurrences in the discussion which follows.

@@

This analysis allows to address the questions raised in Section \ref{sec:suplabelling}:

\textbf{Do label frequencies change over time?} -- In both projects the frequencies generally decreased with age. 
However, there are variations within our NFR labels. In MySQL, \emph{usability} and \emph{efficiency} do not appear very often in topics. 
A proportionately smaller number of commits addressed these NFRs.
As discussed in the previous section, certain peaks in topic numbers coincide with a particular emphasis from the development team on issues such as new releases or bug fixes.
This suggests that maintenance activity is not necessarily strictly decreasing with time, but rather episodic and responsive to outside stimuli. 
In MaxDB, we can observe that \emph{Maintainability} topics became more prevalent as MaxDB matures. 
This is likely due to our analysis timeframe for MaxDB being shorter than the timeframe for the MySQL product.

\textbf{Do the different projects differ in their relative topic interest?} -- 
Yes. MySQL 3.23 had proportionally more
topics labelled \emph{functionality}, while MaxDB had proportionally more
\emph{efficiency} related topics. MaxDB was a very mature release `donated' to the open-source community. 
MySQL, on the other hand, was in its relative infancy. 
Security problems were more common (security is a component of `functionality' in the ISO 9126 model). 
In both cases \emph{portability} was a constant maintenance concern and was prevalent throughout the entire lifetime of the projects.

>>>>

% AH: Topics take a long to read
% AH: Project Dashboards, summarization, software process recovery

Concrete applications of topics and \emph{labelled topic extraction}
range from project dashboards to annotating software artifacts with
NFR-related tags such as revisions and bug reports. 
Project dashboard are typically employed by managers and are used to
provide quick summaries of a the efforts put into a software project,
labelled topic extraction would allow managers to track efforts
related to topics, specifically NFRs.
These techniques also allow the annotation of bug
reports and other software artifacts with the NFRs, such as
portability that they relate to, this would enable querying of bug
reports and artifacts by relevant NFRs. 
For instance one could look
for portability relevant revisions and bug reports.

@@

\subsubsection{The high-level labels}

%\noindent \textbf{The high-level labels} ---

To facilitate cross-project comparison, we used a taxonomy of NFRs. This taxonomy is based on the ISO quality model, ISO9126~\cite{iso9126}. 
ISO9126 describes six high-level non-functional requirements (NFRs), maintainability, functionality, portability, efficiency, usability, and reliability. 
There is some debate about the terms in this model. 
However, ISO9126 is ``an international standard and thus provides an internationally accepted terminology for software quality \cite[p. 58]{Boegh2008},'' that is sufficient for the purposes of this research.  
We claim that these NFRs are maintenance concerns (to varying degrees) in all software projects, and are therefore well-suited to comparison between projects.

%\noindent \textbf{Creating a validation corpus} --- 
\subsubsection{Creating a validation corpus}
To evaluate both unsupervised and supervised classification, we created a validation set of manually labelled topics. For MySQL 3.23 and MaxDB 7.500, we annotated each extracted topic in each period with the same NFR labels as described in the preceding paragraph. 
We looked at each period's topics, and assessed what the data --- consisting of the frequency-weighted word lists and messages --- suggested was the topic for that period. 
We were able to pinpoint the appropriate label using auxiliary information as well, such as the actual revisions and files that were related to the topic being annotated.
For example, for the MaxDB topic consisting of a message ``exit() only used in non NPTL LINUX Versions'', we tagged that topic \emph{portability}. 
%\todo[inline]{this confused people. Is it truly unsupervised? Yes, but it needs to be clearer.}
%We compared against this data-set, but we also used the data-set for our supervised machine learning based topic classification. 
% A topic may have more than one matching keyword. % ok what?
%\todo[inline]{inter-rater reliability }
% \todo[inline]{There should be a section here which is about manual annotation. It is really important and it is required by both analyses. Plus we can tell them how.}

@@

%\todo[inline]{NEIL: abram, i got rid of the letter-grade thing because I'm worried people latch on to the idea we are C- students. I feel anything better than random is our criteria here}
ROC values provide a score %, similar to school letter-grades (A is 0.9, C is 0.6),
 reflecting how well a particular learner performed for the given data. 
ROC maps to the more familiar concepts of precision/sensitivity and recall/specificity: it plots the true positive rate (sensitivity) versus the false positive rate (1 - specificity). 
A perfect learner has a ROC value of 1.0, reflecting perfect recall and precision. 
A ROC result of 0.5 would be equivalent to a random learner (that is, issuing as many false positives as true positives). 
The %area under the -- NE confusing to have area again ... and not importantana
ROC of a classifier is equivalent to the probability that the classifier will rank a randomly chosen positive instance higher than a randomly chosen negative instance.

We consider our labelling classifiers acceptable if they outperform a random classifier (0.5).

\subsection{Unsupervised labelling}
\label{sec:unsuplabelling}

In this section we describe how to label topics based on dictionaries
mined from sources external to the projects themselves.

@@

\subsubsection{Generating word lists}

%\noindent\textbf{Generating word lists} ---

In order to automatically label each topic with an NFR from ISO9126, we associate each NFR with a list
of keywords (`word-lists', in our parlance). These word-lists were determined a priori and were not extracted from the projects themselves.
We intersected the words of the topics and the words of our word-lists.
We `labeled' a topic if any of its words matched any of the word-list's words.
A topic could match more than one NFR.
We used several different sets of word-lists for comparison, which we refer to as \textsf{exp1, exp2, exp3} in the text which follows.

@@

\subsubsection{Automatic Labelled Topic Extraction}

%\todo[inline]{NEIL: this could be a subsubsection}
%\todo[inline]{* the hierarchy of tags was not clear}
%\noindent \textbf{Generating labels automatically} ---

Using our three word-lists (\textsf{exp1}, \textsf{exp2}, \textsf{exp3}), we labeled our topics with an NFR where there was a match between a word in the list and the same word somewhere in the distribution of words that constitute the topic.
A \emph{named topic} is a topic with a matching NFR label. Unnamed topics occur where this no such match, which may indicate either a lack of precision, or simply that this topic is not associated with non-functional requirements.
All experiments were run on MaxDB 7.500 and MySQL 3.23 data. LDA
extracted 20 topics per period for each.
% Is this good enough
This labelling is unsupervised because the corpus is not derived from 
the project being labelled, and we did not label the project's topics
ourselves for a training set. The theory behind this technique is that
because most software often addresses similar issues, we can use the
domain knowledge of software to label relevant topics.
%\todo[inline]{Neil does this explain the unsupervised aspect well enough?}

@@

\subsubsection{Analysis of the unsupervised labelling} %--- %Based
                                %on the labels, and our manual topic
                                %labelling, 
%\noindent \textbf{Analysis of the unsupervised labelling} --- %Based on the labels, and our manual topic labelling, 
%we compared the results of the unsupervised word matching approach. 
For each quality we assessed whether unsupervised labels matched the manual annotations. %To be clear, the manual annotations were not used to train the labeling process.
As described in \ref{sec:wordlist} we used both ROC and F-1 measures to evaluate the performance of the classification.
Figure \ref{fig:maxdb-unsup-results} shows our ROC results for MaxDB and MySQL. We describe F-1 results in the text below.

@@

\subsubsection{Analysis of the supervised labelling}
% \label{sec:suplabelling}
Because our data-set was of word counts we expected Bayesian techniques, often used in spam filtering, to perform well. 
We tried other learners that WEKA~\cite{weka09} provides: rule learners, decision tree learners, vector space learners, and support vector machines.  
Figure \ref{fig:best-learn-per-tag} shows the performance of the best performing learner per label. 
 We considered the best learner for a label to be the learner which
 had the highest ROC value for that label. 
The best learner is important because one uses a single learner per label.
%Figure \ref{fig:best-learn-per-tag} uses the ZeroR learner as a baseline, a common practice in machine learning,
%as ZeroR naively chooses the largest category all of the time. 
%ZeroR occasionally outperforms other learners; for labels which are not as common, this is to be expected because any miscategorization will hurt accuracy. 
%This is why we use ROC values, instead of accuracy, as they can better represent performance on labels which are not applicable to the majority of samples.

@@

\section{Understanding software maintenance activities} 
\label{sec:analysis}
% \todo[inline]{talk about this topic in one section}
% \todo[inline]{reference project name Nomen more}
%In order to test  our ROC values produced acceptable results, we now turn to
%\emph{applying} those techniques to understand software maintenance
%activities. 
We evaluated two research questions using the ground truth data that we
annotated by hand:
\begin{enumerate}
\item Do label frequencies change over time? Is a certain quality of more interest at one point in the life-cycle than some other? 
\item  Do the different projects differ in their relative topic interest? Is a particular quality more important to one project than the other projects?  
\end{enumerate}

% There are several possible applications of \textsf{Nomen}'s techniques. 
% For outside users of the software, \textsf{Nomen} can provide insight into that company's development practices. 
% Are they focused on usability, for example?
% Within the organization \textsf{Nomen} acts as a historical sanity check. 
% Confirmation bias is the difference between what people think happened and what actually happened. 
% Some software releases, for instance, are poorly received since they are quite buggy. 
% \textsf{Nomen} would help explain why this was the case, since it can identify NFRs which are dominating developer attention.
%\todo[inline]{does this section use supervised learning, and if so what ones?}

\subsection{Timeline analysis}
Figures \ref{fig:mysql-timeline} and \ref{fig:maxdb-timeline} show the
temporal patterns of label frequencies based on the manually annotated
topics, although this visualization can be generated from the
results of labelled topic extraction.
There are two measures represented. 
One, the relative frequency, shown in the grey boxes, represents the number of labels with that NFR in that period, 
relative to the maximum number of topics assigned to that NFR. 
The second, absolute frequency, compares the number of topics labelled with that NFR per period 
relative to the maximum number of labelled topics overall, for that project. 
The topmost row in each diagram is reserved for historical events of note. 
We refer to these occurrences in the discussion which follows.

@@

%This analysis allows to address the questions raised in Section \ref{sec:suplabelling}:
This analysis allows to address the questions raised in Section \ref{sec:analysis}:

%\textbf{Do label frequencies change over time?} -- 
\subsubsection{Do label frequencies change over time?}
In both projects the frequencies generally decreased with age. 
However, there are variations within our NFR labels. In MySQL, \emph{usability} and \emph{efficiency} do not appear very often in topics. 
A proportionately smaller number of commits addressed these NFRs.
As discussed in the previous section, certain peaks in topic numbers coincide with a particular emphasis from the development team on issues such as new releases or bug fixes.
This suggests that maintenance activity is not necessarily strictly decreasing with time, but rather episodic and responsive to outside stimuli. 
In MaxDB, we can observe that \emph{Maintainability} topics became more prevalent as MaxDB matures. 
This is likely due to our analysis timeframe for MaxDB being shorter than the timeframe for the MySQL product.

%\textbf{Do the different projects differ in their relative topic
%interest?} --

\subsubsection{Do the different projects differ in their relative
  topic interest?}

Yes. MySQL 3.23 had proportionally more
topics labelled \emph{functionality}, while MaxDB had proportionally more
\emph{efficiency} related topics. MaxDB was a very mature release `donated' to the open-source community. 
MySQL, on the other hand, was in its relative infancy. 
Security problems were more common (security is a component of `functionality' in the ISO 9126 model). 
In both cases \emph{portability} was a constant maintenance concern and was prevalent throughout the entire lifetime of the projects.
